---
title: "Untitled"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
options(stringsAsFactors = F)
```

#Creating folder structure

```{r}
#Terminal commands to created directory structure.

cd /home/jaume/Projects/BreastMapper/Data/
mkdir GSE25066
cd GSE25066/
mkdir Data
mkdir Quality
mkdir Scripts
```


#Download data from GEO in terminal

```{r}
#Terminal commands to download the data.

cd /media/data/jaume/BreastCancer/Data
mkdir GSE25066
cd GSE25066
mkdir Data
cd /media/data/jaume/BreastCancer/Data/GSE25066/Data/
wget https://ftp.ncbi.nlm.nih.gov/geo/series/GSE25nnn/GSE25066/suppl/GSE25066_RAW.tar
wget https://ftp.ncbi.nlm.nih.gov/geo/series/GSE25nnn/GSE25066/soft/GSE25066_family.soft.gz
tar -xf GSE25066_RAW.tar
gunzip *.CEL.gz 
#(Unzip all gz files in a directory)
```

#Loading libraries and functions

```{r}
#Loading libraries.

library(affy)
library(GEOquery)
library(devtools)
library(DupChecker)
library(tools)
library(arrayQualityMetrics)
library(vctrs)
library(frma)
library(a4Preproc)
library(genefu)

#Sourcing functions.

source("/home/jaume/Projects/BreastMapper/Functions/Functions_Breast.R")
```

#GSE25066

#Loading data.

```{r}
"/media/data/jaume/BreastCancer/Data/GSE25066/Data/"
"/media/data/jaume/BreastCancer/Data/GSE25066/Data/"
GEO_dat  <- getGEO(filename= "/media/data/jaume/BreastCancer/Data/GSE25066/Data/GSE25066_family.soft.gz")
```

#Getting information.

```{r}
platforms <- unlist(lapply(GPLList(GEO_dat),function(x) Meta(x)$geo_accession)) 
pData_GEO_dat <- get_pData_table(GEO_dat,platforms[[1]])
rownames(pData_GEO_dat) <- pData_GEO_dat$geo_accession
View(pData_GEO_dat)


pCh_Status <- rep("T",nrow(pData_GEO_dat))
#pCh_Status <- pData_GEO_dat$title
#pCh_Status[grepl("tumor",pCh_Status)] <- "T"
#pCh_Status[grepl("Normal",pCh_Status)] <- "NT"

pCh_Age <- pData_GEO_dat$pCh_age_years

#pCh_MS_By_DS <- pData_GEO_dat$pCh_subtypeihc

#pCh_Histotlogy <- pData_GEO_dat$pCh_histology


#pCh_Status <- rep("T",nrow(pData_GEO_dat))

pCh_T_Code <- pData_GEO_dat$pCh_clinical_t_stage

pCh_N_Code <- pData_GEO_dat$pCh_clinical_nodal_status

#pCh_M_Code <- pData_GEO_dat$pCh_m_stage

#a_temp <- pData_GEO_dat$pCh_time_to_relapse_.years.
#b_temp <- pData_GEO_dat$pCh_time_to_metastasis_.years.
#b_temp[!is.na(a_temp)] <- a_temp[!is.na(a_temp)]
#c_temp <- pData_GEO_dat$pCh_follow_up_duration_.years.
#c_temp[!is.na(b_temp)] <- b_temp[!is.na(b_temp)]

#pCh_DFS_T <- c_temp

#a_temp <- pData_GEO_dat$pCh_regional_relapse
#b_temp <- pData_GEO_dat$pCh_event_metastasis
#b_temp[a_temp=="1"] <- 1

#pCh_DFS_E <- b_temp

pCh_Node <- pData_GEO_dat$pCh_clinical_nodal_status
pCh_Node[pCh_Node == "N0"] <- "0"
pCh_Node[pCh_Node == "N1"] <- "1"
pCh_Node[pCh_Node == "N2"] <- "1"
pCh_Node[pCh_Node == "N3"] <- "1"

#pCh_Size <-  pData_GEO_dat$pCh_size_.bin.
  
pCh_ER_Status <- pData_GEO_dat$pCh_er_status_ihc
pCh_ER_Status[pCh_ER_Status == "P"] <- "1"
pCh_ER_Status[pCh_ER_Status == "N"] <- "0"
pCh_ER_Status[pCh_ER_Status == "NA"] <- NA

pCh_PR_Status <- pData_GEO_dat$pCh_pr_status_ihc
pCh_PR_Status[pCh_PR_Status == "P"] <- "1"
pCh_PR_Status[pCh_PR_Status == "N"] <- "0"
pCh_PR_Status[pCh_PR_Status == "NA"] <- NA

pCh_ERBB2_Status <- pData_GEO_dat$pCh_erbb2_status
pCh_ERBB2_Status[pCh_ERBB2_Status == "N"] <- "0"
pCh_ERBB2_Status[pCh_ERBB2_Status == "P"] <- "1"
#pCh_ERBB2_Status[pCh_ERBB2_Status == "--"] <- NA

#pCh_P53 <- pData_GEO_dat$pCh_p53_ihc

#pCh_KI67 <- pData_GEO_dat$pCh_ki67_ihc

pCh_HER2_Status <- pData_GEO_dat$pCh_her2_status
pCh_HER2_Status[pCh_HER2_Status == "P"] <- "1"
pCh_HER2_Status[pCh_HER2_Status == "N"] <- "0"
pCh_HER2_Status[pCh_HER2_Status == "NA"] <- NA

#pCh_Treatment <- pData_GEO_dat$pCh_endocrine_therapy

pCh_Grade <- pData_GEO_dat$pCh_grade

#pCh_HER2_Fish <- pData_GEO_dat$pCh_her2fish

pCh_DFS_T <- pData_GEO_dat$pCh_drfs_even_time_years
#pCh_DFS_T[pCh_DFS_T == "--"] <- NA
pCh_DFS_T <- as.numeric(pCh_DFS_T)*12

pCh_DFS_E <- pData_GEO_dat$pCh_drfs_1_event_0_censored
#pCh_DFS_E[pCh_DFS_E == "--"] <- NA

#pCh_Death_T <- pData_GEO_dat$pCh_t.os
#pCh_Death_T <- as.numeric(pCh_Death_T)*12
#pCh_Death_T <- as.numeric(pCh_Death_T)/days_to_months

#pCh_Death_E <- pData_GEO_dat$pCh_e.os

pCh_MS_By_DS <- pData_GEO_dat$pCh_pam50_class

#Merging data

vars <- ls()[grepl("pCh_",ls())]
vars <- vars[vars != "pCh_to_Num_and_Fac"]
vars <- vars[vars != "apply_NA_To_Undet_To_pCh_Cols"]

eval(parse(text = paste("pData_GEO_ft <- data.frame(",paste(vars,collapse = ","),")",sep="")))
rownames(pData_GEO_ft) <- rownames(pData_GEO_dat)
View(pData_GEO_ft)

#Removing cel files generated with the B platform.

files_temp <- dir("/media/data/jaume/BreastCancer/Data/GSE25066/Data/",full.names = T)
file.remove(files_temp[grepl("soft.gz",files_temp)])
file.remove(files_temp[grepl("_RAW.tar",files_temp)])


#Obtaining md5 for each .CEL file.

datafile <- buildFileTable(rootDir="/media/data/jaume/BreastCancer/Data/GSE25066/Data")
result <-validateFile(datafile)
result <- result$filetable
rownames(result) <- gsub("_.*","",gsub(".CEL","",gsub(".*/Data/","",result$file)))
result$pCh_samples_to_df <- rownames(result)
result <- result[,-1]
result <- result[,c(3,1,2)]
colnames(result) <- c("pCh_samples_to_df","pCh_directory_to_df","pCh_md5_to_df")
df_out <- result

#Merging

pData_GEO_ft <- merge(pData_GEO_ft,df_out,by = "row.names")
colnames(pData_GEO_ft)[1] <- "pCh_Sample_Name"
rownames(pData_GEO_ft) <- pData_GEO_ft[,1]
View(pData_GEO_ft)

directory_files_temp <- dir("/media/data/jaume/BreastCancer/Data/GSE25066/Data",full.names = T)
directory_files_temp_to_delete <- directory_files_temp[grepl(".CEL.md5$",directory_files_temp)]
for(i in 1:length(directory_files_temp_to_delete)){
  file.remove(directory_files_temp_to_delete[i])
}
```

Reading files.

```{r}
directory_files_temp <- dir("/media/data/jaume/BreastCancer/Data/GSE25066/Data",full.names = T)
eData_GEO <- ReadAffy(filenames=directory_files_temp)
cols_eDat <- colnames(eData_GEO)
cols_eDat <- gsub("_.*","",gsub(".CEL","",cols_eDat))
pData_GEO_ft <- pData_GEO_ft[cols_eDat,]
```

Generating expressionset

```{r}
library("affy")
library("affycoretools")
GEO_Eset_Raw <- affy::rma(eData_GEO,background = FALSE,normalize = FALSE)
colnames(GEO_Eset_Raw)
colnames(GEO_Eset_Raw) <- gsub("_.*","",gsub(".CEL","",colnames(GEO_Eset_Raw)))
pData(GEO_Eset_Raw) <- pData_GEO_ft[colnames(GEO_Eset_Raw),]
exprs(GEO_Eset_Raw)
```

Running array quality analyses using arrayQualityMetrics

```{r}
#source("H:/Tesis_Limpio/Funciones/Calidad_Datos_Sample_Network.R")
arrayQualityMetrics_All_Groups(GEO_Eset_Raw,"/home/jaume/Projects/BreastMapper/Data/GSE25066/Quality/",var = "pCh_Status",do.logtransform = FALSE) #Already tranformed to log scale.
                               
Tum_Data_QC  <- get_outliers_info(get(load(file="/home/jaume/Projects/BreastMapper/Data/GSE25066/Quality/Results/Pre_Norm/T/Raw_array_quality_metrics_Results.Rda")))
nrow(Tum_Data_QC)

delete_tumors <- names(which(rowSums(Tum_Data_QC[,c(1,3,5)]) == 3))
```

Background correction and normalization using RMA

```{r}
eData_GEO_filt <- eData_GEO[,!gsub("_.*","",gsub(".CEL","",colnames(eData_GEO@assayData$exprs))) %in% delete_tumors]
GEO_Eset_Norm <- affy::rma(eData_GEO_filt,background = TRUE,normalize = TRUE)
colnames(GEO_Eset_Norm) <- gsub("_.*","",gsub(".CEL","",colnames(GEO_Eset_Norm)))
pData(GEO_Eset_Norm) <- pData_GEO_ft[colnames(GEO_Eset_Norm),]
dim(GEO_Eset_Norm)

library(a4Preproc)
library(hgu133plus2.db)

GEO_Eset_Norm <- addGeneInfo(GEO_Eset_Norm, annotationLibrary = NULL)
annotation_temp <- featureData(GEO_Eset_Norm)@data
annotation_temp$probe <- rownames(annotation_temp)
colnames(annotation_temp) <- c("EntrezGene.ID","ENSEMBLID","Gene.Symbol","Description","probe")
```

Background correction and normalization using fRMA

```{r}
eData_GEO_filt <- eData_GEO[,!gsub("_.*","",gsub(".CEL","",colnames(eData_GEO@assayData$exprs))) %in% delete_tumors]
GEO_Eset_Norm_frma <- frma(eData_GEO_filt)
exprs(GEO_Eset_Norm_frma)
colnames(GEO_Eset_Norm_frma) <- gsub("_.*","",gsub(".CEL","",colnames(GEO_Eset_Norm_frma)))
pData(GEO_Eset_Norm_frma) <- pData_GEO_ft[colnames(GEO_Eset_Norm_frma),]
dim(GEO_Eset_Norm_frma)

library(a4Preproc)
library(hgu133plus2.db)

GEO_Eset_Norm_frma <- addGeneInfo(GEO_Eset_Norm_frma, annotationLibrary = NULL)
annotation_temp_frma <- featureData(GEO_Eset_Norm_frma)@data
annotation_temp_frma$probe <- rownames(GEO_Eset_Norm_frma)
colnames(annotation_temp_frma) <- c("EntrezGene.ID","ENSEMBLID","Gene.Symbol","Description","probe")
```

Molecular classifications.

```{r}
library(genefu)

pam50 <- molecular.subtyping(sbt.model = "pam50", data = t(exprs(GEO_Eset_Norm)),annot = annotation_temp, do.mapping = TRUE)
pData(GEO_Eset_Norm)$pCh_pam50 <- as.character(pam50$subtype)
pData(GEO_Eset_Norm)$pCh_Batch <- rep("GSE25066",ncol(GEO_Eset_Norm))

pam50_frma <- molecular.subtyping(sbt.model = "pam50", data = t(exprs(GEO_Eset_Norm_frma)),annot = annotation_temp, do.mapping = TRUE)
pData(GEO_Eset_Norm_frma)$pam50_frma <- as.character(pam50_frma$subtype)
pData(GEO_Eset_Norm_frma)$pCh_Batch <- rep("GSE25066",ncol(GEO_Eset_Norm_frma))

table(pData(GEO_Eset_Norm)$pCh_pam50  == pData(GEO_Eset_Norm_frma)$pam50_frma)
```

```{r}
library(survcomp)
library(ggplot2)
list_data_norm <- list(GEO_Eset_Norm,GEO_Eset_Norm_frma)
save(file = "/home/jaume/Projects/BreastMapper/Data/GSE25066/Data/GSE25066_Norm.Rda",list_data_norm)
GEO_Eset_Norm <- get(load(file = "/home/jaume/Projects/BreastMapper/Data/GSE25066/Data/GSE25066_Norm.Rda"))
table(exprs(GEO_Eset_Norm[[1]]) == exprs(GEO_Eset_Norm[[2]]))
```


